---
title: "Methylation CTP validation EDA: Wong2018 ASD methylation"
author: "Chloe Yap"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: hide
---

**To run**: Beforehand:

    module load pandoc

In R:

```{r, eval = FALSE}
setwd("~/shared-gandalm/brain_CTP/Scripts/methylation/analysis/combine_methods/ctp_validation")
rmarkdown::render("ctp_validation_Wong2018_ASDmethylation.Rmd", "html_document")
```

# Set-up

## Directories and libraries

```{r, message = F, warning = F}

library(tidyverse)
library(rmarkdown)    # You need this library to run this template.
library(epuRate)      # Install with remotes::install_github("holtzy/epuRate", force=TRUE) - use this instead of devtools::install_github()
library(data.table)
library(DT)
library(tidyverse)
library(dplyr)
library(reshape2)
library(uwot) # for umap
library(methylCC)
library(lme4)
library(compositions)
library(zCompositions)
library(mixOmics)
source("~/project-gandalm/software/ANCOM-v2.1/scripts/ancom_v2.1.R")
library(ALDEx2)
library(factoextra)
library(ggplot2)
library(RColorBrewer)
library(viridis)
library(ggpubr)
library(GGally)
library(lattice)
library(gridExtra)

# BiocManager::install("M3C")
# library(M3C)

```

```{r}

filen <- "KCL_R01MH094714_ASD_Illumina450K_PFC"
out_dir <- "~/shared-gandalm/brain_CTP/Data/methylation/ASD_methylation_brain/analysis"

# Cell-type number
cellnum <- 9
cellnum <- 7

write_out = TRUE

# Read in reference object
# - use chisq statistic, 5 cell-types, bonferroni p<0.001, marker contributing >=50% of chisq stat, take top 200
#ref_dir <- "~/shared-gandalm/brain_CTP/Data/reference_cell_profile/Luo2020/Luo2020_chisq_dmr_cell5_bonf001_contr50pc_top200.rds"
# - use extreme_methylation_sites.R output
#ref_dir <- "~/shared-gandalm/brain_CTP/Data/reference_cell_profile/Luo2020/Luo2020_extreme17_dmr_low0.3_high0.7.rds"
# - use chisq statistic, 5 cell-types, bonferroni p<0.001, marker contributing >=50% of chisq stat, take top 200
#ref_dir <- "~/shared-gandalm/brain_CTP/Data/reference_cell_profile/Luo2020/Luo2020_chisq_dmr_cell5_bonf001_contr50pc_top200.rds"
#ref_dir <- "~/shared-gandalm/brain_CTP/Data/reference_cell_profile/Luo2020/Luo2020_chisq_dmr_ilmn450kepic_aggto100bp_cell7_bonf001_contr50pc_top200.rds"
#ref <- readRDS(ref_dir)

# Pheno
pheno_dir <- "~/shared-gandalm/brain_CTP/Data/pheno/ASD_methylation_brain/Wong_ASD_Brain_Methylation_SuppTable1_pheno.csv"

# methylCC
# - use Luo et al. reference with DMRs identified from 450K overlap 
#saveRDS(methylcc_out, paste(out_dir, "/", filen, "_aut_mask_methylcc_out_n200_p1e-6.rds", sep = ""))
# - use 450K dlpfc guintivano reference (+/- randomisation of sample order)
#saveRDS(methylcc_out, paste(out_dir, "/", filen, "_aut_mask_methylcc_out_n200_dlpfc_450k_guintivano_hpc.rds", sep = ""))
# - use Luo et al. reference with DMRs pre-identified by Luo et al.
#saveRDS(methylcc_out, paste(out_dir, "/", filen, "_aut_mask_predmr_ilmn450k_celltype_na5_methylcc_dmr_n200_p1e_1.rds", sep = ""))
# - use extreme_methylation_sites.R output
#saveRDS(methylcc_out, paste(out_dir, "/", filen, "_aut_mask_extreme17_dmr_low0.3_high0.7.rds", sep = ""))
# - use chisq statistic, 5 cell-types, bonferroni p<0.001, marker contributing >=50% of chisq stat, take top 200
#saveRDS(methylcc_out, paste(out_dir, "/", filen, "_aut_mask_chisq_dmr_cell5_bonf001_contr50pc_top200.rds", sep = ""))
#methylcc_dir <- paste(out_dir, "/", filen, "_aut_mask_chisq_dmr_cell5_bonf001_contr50pc_top200.rds", sep = "")
filen0 <- "KCL_R01MH094714_ASD_Illumina450K_PFC"
# methylcc_dir <- paste(out_dir, "/", filen0, "_aut_mask_chisq_dmr_cell5_bonf001_contr50pc_top200_aggpostmcc.txt", sep = "")
#methylcc_dir <- paste(out_dir, "/", filen0, "_aut_mask_chisq_dmr_ilmn450kepic_aggto100bp_cell7_bonf001_contr50pc_top200_aggpostmcc.txt", sep = "")

if (cellnum == 9) {
  methylcc_dir <- paste(out_dir, "/", filen0, "_aut_mask_extremes_dmr_ilmn450kepic_aggto100bp_c10_cell9_split6040all_aggpostmcc.txt", sep = "")
} else if (cellnum == 7) {
  methylcc_dir <- paste(out_dir, "/", filen0, "_aut_mask_extremes_dmr_ilmn450kepic_aggto100bp_c10_cell7_split6040all_aggpostmcc.txt", sep = "")
}

# methylCC with NeuN+/- data (Guintivano DLPFC)
methylcc_neun_dir <- "~/shared-gandalm/brain_CTP/Data/methylation/ASD_methylation_brain/analysis/KCL_R01MH094714_ASD_Illumina450K_PFC_aut_mask_methylcc_out_n200_dlpfc_450k_guintivano_hpc.rds"

# eBayes + Houseman
houseman_dir <- "~/shared-gandalm/brain_CTP/Data/methylation/ASD_methylation_brain/analysis/KCL_R01MH094714_ASD_Illumina450K_PFC_ref450K_toast_houseman_ls.rds"

# sequencing with extremes + Houseman
houseman_seq_dir <- "~/shared-gandalm/brain_CTP/Data/methylation/ASD_methylation_brain/analysis/KCL_R01MH094714_ASD_Illumina450K_PFC_Luo2020_extremes_dmr_ilmn450kepic_aggto100bp_c10_cell7_split6040all_houseman_ls.rds"

# celfie
celfie_dir <-  "~/shared-gandalm/brain_CTP/Data/methylation/ASD_methylation_brain/analysis/celfie/KCL_R01MH094714_ASD_Illumina450K_PFC_celfie.out/1_tissue_proportions.txt"
celfie_label_dir <- "~/shared-gandalm/brain_CTP/Data/methylation/ASD_methylation_brain/analysis/celfie/KCL_R01MH094714_ASD_Illumina450K_PFC_celfie.in"

# ReFACToR
#refactor_dir <- "~/shared-gandalm/brain_CTP/Data/methylation/ASD_methylation_brain/analysis/KCL_R01MH094714_ASD_Illumina450K_PFC_k8_cov.refactor.components.txt"

# smartSV
smartsva_dir <- "~/shared-gandalm/brain_CTP/Data/methylation/ASD_methylation_brain/analysis/ASD_methylation_brain_SmartSVA.rds"

# VAE/MethylNet
vae_dir <- "~/shared-gandalm/brain_CTP/Data/integration/methylnet/combined/full_output_latent.csv"

```

## Read-in data

```{r}

# Pheno
pheno <- read.csv(pheno_dir, header = T, as.is = T)
colnames(pheno)[which(colnames(pheno) == "SampleName")] <- "IID"
colnames(pheno)[which(colnames(pheno) == "Age")] <- "age"
colnames(pheno)[which(colnames(pheno) == "Sex")] <- "sex"
pheno$Batch <- as.factor(pheno$Batch)
pheno$group <- ifelse(pheno$Diagnosis == "CTL", "CTL", "ASD")

# methylCC
methylcc_ctp <- read.delim(methylcc_dir)
ind <- grep("NonN", colnames(methylcc_ctp))
foo <- colnames(methylcc_ctp)[ind]
foo2 <- unlist(lapply(strsplit(gsub("NonN_", "", foo), "_"), function(x) x[1]))
colnames(methylcc_ctp)[ind] <- foo2
colnames(methylcc_ctp)[2:ncol(methylcc_ctp)] <- paste("mcc_", colnames(methylcc_ctp)[2:ncol(methylcc_ctp)], sep = "")
# methylcc <- readRDS(methylcc_dir)
# methylcc_ctp <- cell_counts(methylcc)
# ctp_sum <- methylcc@summary
# ctp_input <- ctp_sum$input
# colSums(ctp_input$zmat)
# # Would use this to mimic Supp Fig 1
# colSums(ref$zmat) # this is the initial starting proportions

# methylCC on Guintivano DLPFC reference (NeuN+/-)
# methylcc_neun <- readRDS(methylcc_neun_dir)
# methylcc_neun_ctp <- cell_counts(methylcc_neun)
# colnames(methylcc_neun_ctp) <- c("mcc_NeuN_neg", "mcc_NeuN_pos")
# ctp_neun_sum <- methylcc_neun@summary
# ctp_neun_input <- ctp_neun_sum$input
# colSums(ctp_neun_input$zmat)

# eBayes + Houseman
houseman.ls <- readRDS(houseman_dir)
# - coefs_sub
coefs_sub <- houseman.ls$coefs_sub
colnames(coefs_sub)[2:ncol(coefs_sub)] <- paste("h_", colnames(coefs_sub)[2:ncol(coefs_sub)], sep = "")
colnames(coefs_sub)[ncol(coefs_sub)] <- "h_error_sub"
coefs_brain <- houseman.ls$coefs_brain
colnames(coefs_brain)[2:ncol(coefs_brain)] <- paste("h_", colnames(coefs_brain)[2:ncol(coefs_brain)], sep = "")
colnames(coefs_brain)[ncol(coefs_brain)] <- "h_error_brain"

# sequencing + Houseman
houseman_seq.ls <- readRDS(houseman_seq_dir)
houseman_seq <- houseman_seq.ls[[1]]
ind <- grep("NonN", colnames(houseman_seq))
foo <- colnames(houseman_seq)[ind]
foo2 <- unlist(lapply(strsplit(gsub("NonN_", "", foo), "_"), function(x) x[1]))
colnames(houseman_seq)[ind] <- foo2
colnames(houseman_seq)[2:ncol(houseman_seq)] <- paste("hseq_", colnames(houseman_seq)[2:ncol(houseman_seq)], sep = "")

# CelFIE
celfie <- fread(celfie_dir)
colnames(celfie) <- c("IID_short", "celfie_Exc", "celfie_Inh", "celfie_Astro", "celfie_Endo", "celfie_Micro", "celfie_Oligo", "celfie_OPC", "unknown1")
iid_short <- data.frame(IID = pheno$IID[which(pheno$RegionID == "BA9")])
iid_short$IID_short <- unlist(lapply(strsplit(iid_short$IID, "_"), function(x) x[1]))
celfie <- inner_join(iid_short, celfie, by = "IID_short")
#celfie_label <- fread(celfie_label_dir)

# ReFACToR
# refactor <- read.table(refactor_dir, header = T, as.is = T)
# colnames(refactor)[1] <- "IID"

# smartSVA
smartsva <- readRDS(smartsva_dir)$KCL_R01MH094714_ASD_Illumina450K_PFC
smartsva <- data.frame(IID = rownames(smartsva), smartsva)

# VAE, MethylNet
vae <- read.csv(vae_dir)
colnames(vae) <- c("IID", paste("VAEe", 0:9, sep = ""))

```

### Join into 1 dataframe

```{r}

#ctp_pheno <- inner_join(data.frame(IID = rownames(methylcc_ctp), methylcc_ctp), pheno, by = "IID")
methylcc_ctp.tmp <- methylcc_ctp
methylcc_ctp.tmp$mcc_Glial <- rowSums(methylcc_ctp.tmp[,-c(grep("IID", colnames(methylcc_ctp.tmp)), grep("Exc", colnames(methylcc_ctp.tmp)), grep("Inh", colnames(methylcc_ctp.tmp)))])
methylcc_ctp.tmp$mcc_Neuron <- rowSums(methylcc_ctp.tmp[,c(grep("Exc", colnames(methylcc_ctp.tmp)), grep("Inh", colnames(methylcc_ctp.tmp)))])
ctp_pheno <- inner_join(methylcc_ctp.tmp, pheno, by = "IID")
#ctp_pheno <- inner_join(ctp_pheno, data.frame(IID = rownames(methylcc_neun_ctp), methylcc_neun_ctp), by = "IID")
ctp_pheno <- inner_join(ctp_pheno, smartsva, by = "IID")

# Houseman with array reference
ctp_pheno <- inner_join(ctp_pheno, coefs_sub, by = "IID")
ctp_pheno <- inner_join(ctp_pheno, coefs_brain, by = "IID")
ctp_pheno$h_Glial <- rowSums(ctp_pheno[,c("h_ASTRO", "h_OPC", "h_OLIGO", "h_ENDO", "h_MONO")])
ctp_pheno$h_Neuron <- rowSums(ctp_pheno[,c("h_GABA", "h_GLU")])

# Houseman with sequencing reference
ctp_pheno <- inner_join(ctp_pheno, houseman_seq, by = "IID")
ctp_pheno$hseq_Glial <- rowSums(ctp_pheno[,c("hseq_Astro", "hseq_Endo", "hseq_Micro", "hseq_Oligo", "hseq_OPC")])
ctp_pheno$hseq_Neuron <- rowSums(ctp_pheno[,c("hseq_Exc", "hseq_Inh")])

# CelFIE
ctp_pheno <- inner_join(ctp_pheno, celfie, by = c("IID"))
ctp_pheno$celfie_Glial <- rowSums(ctp_pheno[,c("celfie_Astro", "celfie_Endo", "celfie_Micro", "celfie_Oligo", "celfie_OPC")])
ctp_pheno$celfie_Neuron <- rowSums(ctp_pheno[,c("celfie_Exc", "celfie_Inh")])

# Choice to exclude Oxford brain bank subset
# - choose not to as lose statistical power ...
# ctp_pheno <- ctp_pheno %>% filter(BrainBank != "Oxford")

ctp_pheno <- left_join(ctp_pheno, vae, by = "IID")

```

### Write output

```{r}

if (write_out == TRUE) {
  write.table(ctp_pheno, paste(out_dir, "/", filen, "_ctp_pheno.txt", sep = ""), col.names = T, row.names = F, quote = F, sep = "\t")
}

```

### Set variables

```{r}

keep_cols <- c("IID", "group", "age", "sex", "Batch", "PMI", "BrainBank")

ctp_cols <- c(colnames(ctp_pheno)[grep("hseq", colnames(ctp_pheno))])
ctp_cols <- ctp_cols[-which(ctp_cols == "hseq_error")]
level2_cols <- ctp_cols[-which(ctp_cols %in% c("hseq_Glial", "hseq_Neuron"))]

glial_rmoligo_n <- colnames(ctp_pheno)[grep("hseq", colnames(ctp_pheno))]
glial_rmoligo_n <- glial_rmoligo_n[-which(glial_rmoligo_n %in% c("hseq_Exc", "hseq_Inh", "hseq_Oligo", "hseq_error", "hseq_Glial", "hseq_Neuron"))]
neuron_n <- c("hseq_Exc", "hseq_Inh")

```

# Check baseline variable differences

- look for confounding
- there is confounding for age, brain bank, batch (when looking at the PFC samples only)
- this means that accounting for covariates might also take out the ASD signal?!

```{r}

# BrainBank
brainbank <- ctp_pheno %>% group_by(group, BrainBank) %>% summarise(n=n()) %>% spread(group, n)
brainbank
chisq.test(brainbank %>% dplyr::select("ASD", "CTL"))

# Batch
batch <- ctp_pheno %>% group_by(group, Batch) %>% summarise(n=n()) %>% spread(group, n)
batch
#chisq.test(batch %>% dplyr::select("ASD", "CTL"))

# Sex
sex <- ctp_pheno %>% group_by(group, sex) %>% summarise(n=n()) %>% spread(group, n)
sex
chisq.test(sex %>% dplyr::select("ASD", "CTL"))

# Age
summary(lm(age ~ group, data = ctp_pheno))
ggplot(ctp_pheno, aes(x = age, colour = group, fill = group)) + geom_histogram() + facet_wrap(~group)

# BrainBank x age x group
ctp_pheno %>% group_by(BrainBank, group) %>% summarise(mean = mean(age), sd = sd(age))

```

# Check association of sSVs and batch effects

```{r}

tmp.long <- ctp_pheno %>% dplyr::select(c("IID", "BrainBank", "Batch", "sex", "age", "group", "hseq_Glial", "sSV1", "sSV2", "sSV3", "sSV4", "sSV5", "sSV6", "sSV7", "sSV8")) %>% melt(id.vars = c("IID", "BrainBank", "Batch", "sex", "age", "group", "hseq_Glial"), variable.name = "sSV_num", value.name = "sSV")

ggplot(tmp.long, aes(x = BrainBank, y = sSV)) + geom_boxplot() + facet_wrap(~ sSV_num, scales = "free_y")
ggplot(tmp.long, aes(x = Batch, y = sSV, colour = Batch)) + geom_boxplot() + facet_wrap(~ sSV_num, scales = "free_y")
ggplot(tmp.long, aes(x = Batch, y = sSV, colour = BrainBank)) + geom_boxplot() + facet_wrap(~ sSV_num, scales = "free_y")
ggplot(tmp.long, aes(x = group, y = sSV, colour = group)) + geom_boxplot() + facet_wrap(~ sSV_num, scales = "free_y")
ggplot(tmp.long, aes(x = sex, y = sSV)) + geom_boxplot() + facet_wrap(~ sSV_num)
ggplot(tmp.long, aes(x = age, y = sSV, colour = BrainBank)) + geom_point() + geom_smooth(method = "lm") + facet_wrap(~ sSV_num)
ggplot(tmp.long, aes(x = hseq_Glial, y = sSV, colour = BrainBank)) + geom_point() + geom_smooth(method = "lm") + facet_wrap(~ sSV_num)

```

# Check association of VAE embeddings and batch effects

```{r}

tmp.long <- ctp_pheno %>% dplyr::select(c("IID", "BrainBank", "Batch", "sex", "age", "group", "hseq_Glial", "VAEe0", "VAEe1", "VAEe2", "VAEe3", "VAEe4", "VAEe5", "VAEe6", "VAEe7", "VAEe8", "VAEe9")) %>% melt(id.vars = c("IID", "BrainBank", "Batch", "sex", "age", "group", "hseq_Glial"), variable.name = "VAEe", value.name = "embedding")

ggplot(tmp.long, aes(x = BrainBank, y = embedding)) + geom_boxplot() + facet_wrap(~ VAEe, scales = "free_y")
ggplot(tmp.long, aes(x = Batch, y = embedding, colour = Batch)) + geom_boxplot() + facet_wrap(~ VAEe, scales = "free_y")
ggplot(tmp.long, aes(x = Batch, y = embedding, colour = BrainBank)) + geom_boxplot() + facet_wrap(~ VAEe, scales = "free_y")
ggplot(tmp.long, aes(x = group, y = embedding, colour = group)) + geom_boxplot() + facet_wrap(~ VAEe, scales = "free_y")
ggplot(tmp.long, aes(x = sex, y = embedding)) + geom_boxplot() + facet_wrap(~ VAEe)
ggplot(tmp.long, aes(x = age, y = embedding, colour = BrainBank)) + geom_point() + geom_smooth(method = "lm") + facet_wrap(~ VAEe)
ggplot(tmp.long, aes(x = hseq_Glial, y = embedding, colour = BrainBank)) + geom_point() + geom_smooth(method = "lm") + facet_wrap(~ VAEe)

```

# Boxplots - raw

Add analysis info on

```{r, fig.width = 8, fig.height = 12}

houseman_seq.long <- melt(houseman_seq, variable.name = "celltype", value.name = "seq_houseman_CTP")
houseman_seq.gg <- ggplot(houseman_seq.long, aes(x = celltype, y = seq_houseman_CTP, colour = celltype)) + geom_boxplot() + theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + ylim(c(0,1))

methylcc_ctp.long <- melt(methylcc_ctp, variable.name = "celltype", value.name = "methylcc_CTP")
methylcc_ctp.gg <- ggplot(methylcc_ctp.long, aes(x = celltype, y = methylcc_CTP, colour = celltype)) + geom_boxplot() + theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + ylim(c(0,1))
#+ theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))

coefs_sub.long <- melt(coefs_sub, variable.name = "celltype", value.name = "eBayes_houseman_CTP") %>% mutate(MatchCell = match(celltype, c("h_GLU", "h_GABA", "h_ASTRO", "h_ENDO", "h_MONO", "h_OLIGO", "h_OPC")))
coefs_sub.gg <- coefs_sub.long %>% mutate(celltype = fct_reorder(celltype, MatchCell)) %>% ggplot(aes(x = celltype, y = eBayes_houseman_CTP, colour = celltype)) + geom_boxplot() + theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + ylim(c(0,1))

coefs_brain.long <- melt(coefs_brain, variable.name = "celltype", value.name = "eBayes_houseman_CTP")
coefs_brain.gg <- ggplot(coefs_brain.long, aes(x = celltype, y = eBayes_houseman_CTP, colour = celltype)) + geom_boxplot() + theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + ylim(c(0,1))

smartsva.long <- melt(smartsva, variable.name = "celltype", value.name = "smartsva_CTP")
smartsva.gg <- ggplot(smartsva.long, aes(x = celltype, y = smartsva_CTP, colour = celltype)) + geom_boxplot()  + theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

vae.long <- melt(vae, id.vars = c("IID"), variable.name = "celltype", value.name = "VAE_embed")
vae.gg <- ggplot(vae.long, aes(x = celltype, y = VAE_embed, colour = celltype)) + geom_boxplot() + theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + ylim(c(0,1))

celfie.long <- melt(celfie, id.vars = c("IID", "IID_short"), variable.name = "celltype", value.name = "CelFIE_CTP")
celfie.gg <- ggplot(celfie.long, aes(x = celltype, y = CelFIE_CTP, colour = celltype)) + geom_boxplot() + theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + ylim(c(0,1))

ggarrange(houseman_seq.gg, methylcc_ctp.gg, coefs_sub.gg, coefs_brain.gg, celfie.gg, smartsva.gg, nrow = 3, ncol = 2)

```

## houseman_seq

```{r}

# - boxplot
hseq_ctp_pheno.tmp <- ctp_pheno %>% dplyr::select(c(all_of(keep_cols), all_of(ctp_cols)))

if (write_out == TRUE) {
  write.table(hseq_ctp_pheno.tmp, paste(out_dir, "/hseq_ctp_pheno.raw.txt", sep = ""), col.names = T, row.names = F, quote = F, sep = "\t")
}

```

```{r}

hseq_ctp_pheno.long <- melt(hseq_ctp_pheno.tmp, measure.vars = c(all_of(ctp_cols)), variable.name = "celltype", value.name = "CTP")
hseq_ctp_pheno.long$Level <- ifelse(hseq_ctp_pheno.long$celltype %in% c("hseq_Neuron", "hseq_Glial"), "Level1", "Level2")
hseq_ctp_pheno.long$Region <- "PFC"
hseq_ctp_pheno.long$celltype <- gsub("hseq_", "", hseq_ctp_pheno.long$celltype)
hseq_ctp_pheno.long$celltype <- unlist(lapply(strsplit(hseq_ctp_pheno.long$celltype, "_"), function(x) x[1]))
hseq_ctp_pheno.long$MatchCell <- match(hseq_ctp_pheno.long$celltype, rev(c("Exc", "Inh", "Astro", "Micro", "Endo", "Oligo", "OPC", "Neuron", "Glial")))
hseq_ctp_pheno.long %>% mutate(celltype = fct_reorder(celltype, MatchCell)) %>%
  ggplot(aes(x=celltype, y = CTP, fill = group)) + 
  geom_boxplot(position = position_dodge(.8),alpha=.8,outlier.size = 0)+ 
  geom_point(position = position_jitterdodge(.1),alpha=.3) +   
    facet_grid(Level~Region, scales='free', space = 'free_y') + coord_flip() + theme_bw() + xlab("")

```

### Adjust for oligodendrocyte proportion

```{r}

# New adjustment which only changes neuronal populations
hseq_ctp_pheno_adjoligo.tmp <- hseq_ctp_pheno.tmp %>%
  mutate(sum_neuro_oligo = hseq_Exc + hseq_Inh + hseq_Oligo) %>%
  mutate(Exc_Inh_ratio = (hseq_Exc/(hseq_Exc + hseq_Inh))) %>%
  mutate(Inh_Exc_ratio = (hseq_Inh/(hseq_Exc + hseq_Inh))) %>%
  mutate(hseq_Exc = Exc_Inh_ratio * sum_neuro_oligo) %>%
  mutate(hseq_Inh = Inh_Exc_ratio * sum_neuro_oligo) %>%
  dplyr::select(-c("hseq_Oligo", "hseq_Neuron", "hseq_Glial"))

# Old adjustment which changed ALL CTPs
# methylcc_ctp_pheno_adjoligo.tmp$sum_nonoligo <- methylcc_ctp_pheno_adjoligo.tmp$Exc + methylcc_ctp_pheno_adjoligo.tmp $Inh + methylcc_ctp_pheno_adjoligo.tmp$NonN_Astro_FGF3R + methylcc_ctp_pheno_adjoligo.tmp$NonN_Micro.Endo_TYROBP
# methylcc_ctp_pheno_adjoligo.tmp$Exc <- methylcc_ctp_pheno_adjoligo.tmp$Exc/methylcc_ctp_pheno_adjoligo.tmp$sum_nonoligo
# methylcc_ctp_pheno_adjoligo.tmp$Inh <- methylcc_ctp_pheno_adjoligo.tmp$Inh/methylcc_ctp_pheno_adjoligo.tmp$sum_nonoligo
# methylcc_ctp_pheno_adjoligo.tmp$NonN_Astro_FGF3R <- methylcc_ctp_pheno_adjoligo.tmp$NonN_Astro_FGF3R/methylcc_ctp_pheno_adjoligo.tmp$sum_nonoligo
# methylcc_ctp_pheno_adjoligo.tmp$NonN_Micro.Endo_TYROBP <- methylcc_ctp_pheno_adjoligo.tmp$NonN_Micro.Endo_TYROBP/methylcc_ctp_pheno_adjoligo.tmp$sum_nonoligo

hseq_ctp_pheno_adjoligo.tmp$hseq_Neuron <- rowSums(hseq_ctp_pheno_adjoligo.tmp[,neuron_n])
hseq_ctp_pheno_adjoligo.tmp$hseq_Glial <- rowSums(hseq_ctp_pheno_adjoligo.tmp[,glial_rmoligo_n])

if (write_out == TRUE) {
  write.table(hseq_ctp_pheno_adjoligo.tmp, paste(out_dir, "/hseq_adjoligo_ctp_pheno.raw.txt", sep = ""), col.names = T, row.names = F, quote = F, sep = "\t")
}

```

```{r}

hseq_ctp_pheno_adjoligo.long <- melt(hseq_ctp_pheno_adjoligo.tmp, measure.vars = c(all_of(neuron_n), all_of(glial_rmoligo_n), "hseq_Neuron", "hseq_Glial"), variable.name = "celltype", value.name = "CTP")
hseq_ctp_pheno_adjoligo.long$Level <- ifelse(hseq_ctp_pheno_adjoligo.long$celltype %in% c("hseq_Neuron", "hseq_Glial"), "Level1", "Level2")
hseq_ctp_pheno_adjoligo.long$Region <- "PFC"
hseq_ctp_pheno_adjoligo.long$celltype <- gsub("hseq_", "", hseq_ctp_pheno_adjoligo.long$celltype)

hseq_ctp_pheno_adjoligo.long %>% mutate(celltype = fct_reorder(celltype, match(celltype, rev(c("Exc", "Inh", "Astro", "Endo", "Micro", "OPC"))))) %>%
  ggplot(aes(x=celltype, y = CTP, fill = group)) + 
  geom_boxplot(position = position_dodge(.8),alpha=.8,outlier.size = 0)+ 
  geom_point(position = position_jitterdodge(.1),alpha=.3) +   
    facet_grid(Level~Region, scales='free', space = 'free_y') + coord_flip() + theme_bw() + xlab("")

```


## methylCC

```{r}

ctp_cols_mcc <- gsub("hseq", "mcc", ctp_cols)
neuron_n_mcc <- gsub("hseq", "mcc", neuron_n)
glial_rmoligo_n_mcc <- gsub("hseq", "mcc", glial_rmoligo_n)

# - boxplot
methylcc_ctp_pheno.tmp <- ctp_pheno %>% dplyr::select(c(all_of(keep_cols), all_of(ctp_cols_mcc)))

if (write_out == TRUE) {
  write.table(methylcc_ctp_pheno.tmp, paste(out_dir, "/methylcc_ctp_pheno.raw.txt", sep = ""), col.names = T, row.names = F, quote = F, sep = "\t")
}

```

```{r}

methylcc_ctp_pheno.long <- melt(methylcc_ctp_pheno.tmp, measure.vars = c(all_of(ctp_cols_mcc)), variable.name = "celltype", value.name = "CTP")
methylcc_ctp_pheno.long$Level <- ifelse(methylcc_ctp_pheno.long$celltype %in% c("mcc_Neuron", "mcc_Glial"), "Level1", "Level2")
methylcc_ctp_pheno.long$Region <- "PFC"
methylcc_ctp_pheno.long$celltype <- gsub("mcc_", "", methylcc_ctp_pheno.long$celltype)

methylcc_ctp_pheno.long %>% mutate(celltype = fct_reorder(celltype, match(celltype, rev(c("Exc", "Inh", "Astro", "Endo", "Micro", "Oligo", "OPC"))))) %>%
ggplot(aes(x=celltype, y = CTP, fill = group)) + 
  geom_boxplot(position = position_dodge(.8),alpha=.8,outlier.size = 0)+ 
  geom_point(position = position_jitterdodge(.1),alpha=.3) +   
    facet_grid(Level~Region, scales='free', space = 'free_y') + coord_flip() + theme_bw() + xlab("")

```

### Adjust for oligodendrocyte proportion

```{r}

# New adjustment which only changes neuronal populations
methylcc_ctp_pheno_adjoligo.tmp <- methylcc_ctp_pheno.tmp %>%
  mutate(sum_neuro_oligo = mcc_Exc + mcc_Inh + mcc_Oligo) %>%
  mutate(Exc_Inh_ratio = (mcc_Exc/(mcc_Exc +mcc_Inh))) %>%
  mutate(Inh_Exc_ratio = (mcc_Inh/(mcc_Exc + mcc_Inh))) %>%
  mutate(mcc_Exc = Exc_Inh_ratio * sum_neuro_oligo) %>%
  mutate(mcc_Inh = Inh_Exc_ratio * sum_neuro_oligo) %>%
  dplyr::select(-c("mcc_Oligo", "mcc_Neuron", "mcc_Glial"))

# Old adjustment which changed ALL CTPs
# methylcc_ctp_pheno_adjoligo.tmp$sum_nonoligo <- methylcc_ctp_pheno_adjoligo.tmp$Exc + methylcc_ctp_pheno_adjoligo.tmp $Inh + methylcc_ctp_pheno_adjoligo.tmp$NonN_Astro_FGF3R + methylcc_ctp_pheno_adjoligo.tmp$NonN_Micro.Endo_TYROBP
# methylcc_ctp_pheno_adjoligo.tmp$Exc <- methylcc_ctp_pheno_adjoligo.tmp$Exc/methylcc_ctp_pheno_adjoligo.tmp$sum_nonoligo
# methylcc_ctp_pheno_adjoligo.tmp$Inh <- methylcc_ctp_pheno_adjoligo.tmp$Inh/methylcc_ctp_pheno_adjoligo.tmp$sum_nonoligo
# methylcc_ctp_pheno_adjoligo.tmp$NonN_Astro_FGF3R <- methylcc_ctp_pheno_adjoligo.tmp$NonN_Astro_FGF3R/methylcc_ctp_pheno_adjoligo.tmp$sum_nonoligo
# methylcc_ctp_pheno_adjoligo.tmp$NonN_Micro.Endo_TYROBP <- methylcc_ctp_pheno_adjoligo.tmp$NonN_Micro.Endo_TYROBP/methylcc_ctp_pheno_adjoligo.tmp$sum_nonoligo

methylcc_ctp_pheno_adjoligo.tmp$mcc_Neuron <- rowSums(methylcc_ctp_pheno_adjoligo.tmp[,neuron_n_mcc])
methylcc_ctp_pheno_adjoligo.tmp$mcc_Glial <- rowSums(methylcc_ctp_pheno_adjoligo.tmp[,glial_rmoligo_n_mcc])

if (write_out == TRUE) {
  write.table(methylcc_ctp_pheno_adjoligo.tmp, paste(out_dir, "/methylcc_adjoligo_ctp_pheno.raw.txt", sep = ""), col.names = T, row.names = F, quote = F, sep = "\t")
}

```

```{r}

methylcc_ctp_pheno_adjoligo.long <- melt(methylcc_ctp_pheno_adjoligo.tmp, measure.vars = c(all_of(neuron_n_mcc), all_of(glial_rmoligo_n_mcc), "mcc_Neuron", "mcc_Glial"), variable.name = "celltype", value.name = "CTP")
methylcc_ctp_pheno_adjoligo.long$Level <- ifelse(methylcc_ctp_pheno_adjoligo.long$celltype %in% c("mcc_Neuron", "mcc_Glial"), "Level1", "Level2")
methylcc_ctp_pheno_adjoligo.long$Region <- "PFC"
methylcc_ctp_pheno_adjoligo.long$celltype <- gsub("mcc_", "", methylcc_ctp_pheno_adjoligo.long$celltype)

methylcc_ctp_pheno_adjoligo.long %>% mutate(celltype = fct_reorder(celltype, match(celltype, rev(c("Exc", "Inh", "Astro", "Endo", "Micro", "OPC"))))) %>%
ggplot(aes(x=celltype, y = CTP, fill = group)) + 
  geom_boxplot(position = position_dodge(.8),alpha=.8,outlier.size = 0)+ 
  geom_point(position = position_jitterdodge(.1),alpha=.3) +   
    facet_grid(Level~Region, scales='free', space = 'free_y') + coord_flip() + theme_bw() + xlab("")

```

## coefs_sub

```{r}

# - boxplot
coefs_sub_ctp_pheno.tmp <- ctp_pheno %>% dplyr::select(c(all_of(keep_cols), "h_GLU", "h_GABA", "h_ASTRO", "h_ENDO", "h_MONO", "h_OPC", "h_OLIGO"))
coefs_sub_ctp_pheno.tmp$neuron_sum <- coefs_sub_ctp_pheno.tmp$h_GABA + coefs_sub_ctp_pheno.tmp$h_GLU
coefs_sub_ctp_pheno.tmp$glia_sum <- coefs_sub_ctp_pheno.tmp$h_ASTRO + coefs_sub_ctp_pheno.tmp$h_OPC + coefs_sub_ctp_pheno.tmp$h_OLIGO + coefs_sub_ctp_pheno.tmp$h_ENDO + coefs_sub_ctp_pheno.tmp$h_MONO

if (write_out == TRUE) {
  write.table(coefs_sub_ctp_pheno.tmp, paste(out_dir, "/coefs_sub_ctp_pheno.raw.txt", sep = ""), col.names = T, row.names = F, quote = F, sep = "\t")
}

```

```{r}

coefs_sub_ctp_pheno.long <- melt(coefs_sub_ctp_pheno.tmp, measure.vars = c("h_GLU", "h_GABA", "h_ASTRO", "h_ENDO", "h_MONO", "h_OPC", "h_OLIGO", "neuron_sum", "glia_sum"), variable.name = "celltype", value.name = "CTP")
coefs_sub_ctp_pheno.long$Level <- ifelse(coefs_sub_ctp_pheno.long$celltype %in% c("neuron_sum", "glia_sum"), "Level1", "Level2")
coefs_sub_ctp_pheno.long$Region <- "PFC"

ggplot(coefs_sub_ctp_pheno.long, aes(x=celltype, y = CTP, fill = group)) + 
  geom_boxplot(position = position_dodge(.8),alpha=.8,outlier.size = 0)+ 
  geom_point(position = position_jitterdodge(.1),alpha=.3) +   
    facet_grid(Level~Region, scales='free', space = 'free_y') + coord_flip() + theme_bw() + xlab("")

```

# Boxplots: clr-transformation (2 ways to handle zeros)

- Justification: does it make sense to run an ANOVA on CTP ~ group, if the CTPs are related? 
- Eg. what does variance analysis look like when the y-variable for each individual is non-independent?

## houseman_seq

### clr-transformation with z-compositions to deal with zeros

- IDs go as columns, CTPs as rows for zCompositions cmultRepl(), and IDs as rows for the clr(?

```{r}

keep_pheno_col <- hseq_ctp_pheno.tmp %>% dplyr::select(all_of(keep_cols))
hseq_ctp_pheno.tmp.clr <- as.matrix(hseq_ctp_pheno.tmp %>% dplyr::select(c(all_of(level2_cols))))

# Apply zCompositions to empirically get the best replacement of zeros
# - it looks like you have to transpose the matrix (with samples as columns) to do this properly (?)
length(which(hseq_ctp_pheno.tmp.clr == 0)) # check there are no 0s
check <- which(hseq_ctp_pheno.tmp.clr == 0, arr.ind = T)

hseq_ctp_pheno.tmp.clr_t <- t(hseq_ctp_pheno.tmp.clr)

if (length(which(hseq_ctp_pheno.tmp.clr == 0)) > 0) {
  
  check <- which(hseq_ctp_pheno.tmp.clr_t == 0, arr.ind = T)
  hseq_ctp_pheno.tmp.clr0_t <- cmultRepl(hseq_ctp_pheno.tmp.clr_t, output = "p-counts")
  hseq_ctp_pheno.tmp.clr0_t[check]

} else {
  
  print("no zcompositions applied as no 0 values")
  hseq_ctp_pheno.tmp.clr0_t <- hseq_ctp_pheno.tmp.clr_t
  
}

# Perform clr-transform
hseq_ctp_pheno.tmp.clr0.tr <- clr(t(hseq_ctp_pheno.tmp.clr_t))
hseq_ctp_pheno.clr.zc <- data.frame(keep_pheno_col, hseq_ctp_pheno.tmp.clr0.tr)

if (write_out == TRUE) {
  write.table(hseq_ctp_pheno.clr.zc, paste(out_dir, "/hseq_ctp_pheno.clr.zc.txt", sep = ""), col.names = T, row.names = F, quote = F, sep = "\t")
}

```

**Make dataframe**

```{r}

hseq_ctp_pheno.clr.zc.long <- melt(hseq_ctp_pheno.clr.zc, measure.vars = c(all_of(level2_cols)), variable.name = "celltype", value.name = "CTP")
#methylcc_ctp_pheno.clr.zc.long$Level <- ifelse(methylcc_ctp_pheno.clr.zc.long$celltype %in% c("neuron_sum", "glia_sum"), "Level1", "Level2")
hseq_ctp_pheno.clr.zc.long$Region <- "PFC"

```

```{r}

ggplot(hseq_ctp_pheno.clr.zc.long, aes(x=celltype, y = CTP, fill = group)) + 
  geom_boxplot(position = position_dodge(.8),alpha=.8,outlier.size = 0)+ 
  geom_point(position = position_jitterdodge(.1),alpha=.3) +
  coord_flip() + theme_bw() + xlab("")
  ggtitle("zCompositions")
    #facet_grid(Region, scales='free', space = 'free_y') + coord_flip() + theme_bw() + xlab("")

```

### clr-transformation with 1e-3 minimum

- Justification: does it make sense to run an ANOVA on CTP ~ group, if the CTPs are related? 
- Eg. what does variance analysis look like when the y-variable for each individual is non-independent?

- N.B. IDs go as columns, CTPs as rows

```{r}

keep_pheno_col <- hseq_ctp_pheno.tmp %>% dplyr::select(all_of(keep_cols))
hseq_ctp_pheno.tmp.clr <- as.matrix(hseq_ctp_pheno.tmp %>% dplyr::select(c(all_of(level2_cols))))

# Make minimum CTP = 1e-3
length(which(hseq_ctp_pheno.tmp.clr <= 1e-3)) # check there are no 0s
hseq_ctp_pheno.tmp.clr0 <- hseq_ctp_pheno.tmp.clr
hseq_ctp_pheno.tmp.clr0[which(hseq_ctp_pheno.tmp.clr0 <= 1e-3)] <- 1e-3

# Perform clr-transform
hseq_ctp_pheno.tmp.clr0.tr <- clr(hseq_ctp_pheno.tmp.clr0)
hseq_ctp_pheno.clr.1e3 <- data.frame(keep_pheno_col, hseq_ctp_pheno.tmp.clr0.tr)

if (write_out == TRUE) {
  write.table(hseq_ctp_pheno.clr.1e3, paste(out_dir, "/hseq_ctp_pheno.clr.1e3.txt", sep = ""), col.names = T, row.names = F, quote = F, sep = "\t")
}

```

**Make dataframe**

```{r}

hseq_ctp_pheno.clr.1e3.long <- melt(hseq_ctp_pheno.clr.1e3, measure.vars = c(all_of(level2_cols)), variable.name = "celltype", value.name = "CTP")
#methylcc_ctp_pheno.clr.zc.long$Level <- ifelse(methylcc_ctp_pheno.clr.zc.long$celltype %in% c("neuron_sum", "glia_sum"), "Level1", "Level2")
hseq_ctp_pheno.clr.1e3.long$Region <- "PFC"

```

```{r}

ggplot(hseq_ctp_pheno.clr.1e3.long, aes(x=celltype, y = CTP, fill = group)) + 
  geom_boxplot(position = position_dodge(.8),alpha=.8,outlier.size = 0)+ 
  geom_point(position = position_jitterdodge(.1),alpha=.3) +
  coord_flip() + theme_bw() + xlab("") +
  ggtitle("Minimum value = 1e-3")
    #facet_grid(Region, scales='free', space = 'free_y') + coord_flip() + theme_bw() + xlab("")

```

#### Adjust for oligodendrocyte proportion

```{r}

keep_pheno_col <- hseq_ctp_pheno_adjoligo.tmp %>% dplyr::select(all_of(keep_cols))
hseq_ctp_pheno_adjoligo.tmp.clr <- as.matrix(hseq_ctp_pheno_adjoligo.tmp %>% dplyr::select(c(all_of(neuron_n), all_of(glial_rmoligo_n))))

# Make minimum CTP = 1e-3
length(which(hseq_ctp_pheno_adjoligo.tmp.clr <= 1e-3)) # check there are no 0s
hseq_ctp_pheno_adjoligo.tmp.clr0 <- hseq_ctp_pheno_adjoligo.tmp.clr
hseq_ctp_pheno_adjoligo.tmp.clr0[which(hseq_ctp_pheno_adjoligo.tmp.clr0 <= 1e-3)] <- 1e-3

# Perform clr-transform
hseq_ctp_pheno_adjoligo.tmp.clr0.tr <- clr(hseq_ctp_pheno_adjoligo.tmp.clr0)
hseq_ctp_pheno_adjoligo.clr.1e3 <- data.frame(keep_pheno_col, hseq_ctp_pheno_adjoligo.tmp.clr0.tr)

if (write_out == TRUE) {
  write.table(hseq_ctp_pheno_adjoligo.clr.1e3, paste(out_dir, "/hseq_adjoligo_ctp_pheno.clr.1e3.txt", sep = ""), col.names = T, row.names = F, quote = F, sep = "\t")
}

```

**Make dataframe**

```{r}

hseq_ctp_pheno_adjoligo.clr.1e3.long <- melt(hseq_ctp_pheno_adjoligo.clr.1e3, measure.vars = c(all_of(neuron_n), all_of(glial_rmoligo_n)), variable.name = "celltype", value.name = "CTP")
#methylcc_ctp_pheno.clr.zc.long$Level <- ifelse(methylcc_ctp_pheno.clr.zc.long$celltype %in% c("neuron_sum", "glia_sum"), "Level1", "Level2")
hseq_ctp_pheno_adjoligo.clr.1e3.long$Region <- "PFC"

```

```{r}

ggplot(hseq_ctp_pheno_adjoligo.clr.1e3.long, aes(x=celltype, y = CTP, fill = group)) + 
  geom_boxplot(position = position_dodge(.8),alpha=.8,outlier.size = 0)+ 
  geom_point(position = position_jitterdodge(.1),alpha=.3) +
  coord_flip() + theme_bw() + xlab("") +
  ggtitle("Minimum value = 1e-3")
    #facet_grid(Region, scales='free', space = 'free_y') + coord_flip() + theme_bw() + xlab("")

```

# Model: per-cell-type, compare raw

```{r}

if (cellnum == 7) {
  
  summary(lm(cbind(hseq_Exc, hseq_Inh, hseq_Astro, hseq_Endo, hseq_Micro, hseq_Oligo, hseq_OPC) ~ group, data = hseq_ctp_pheno.tmp))

  summary(lm(cbind(hseq_Exc, hseq_Inh, hseq_Astro, hseq_Endo, hseq_Micro, hseq_Oligo, hseq_OPC) ~ group + age + sex + Batch + PMI, data = hseq_ctp_pheno.tmp))

} else if (cellnum == 9) {
  
  summary(lm(cbind(Exc_upper, Exc_lower, Inh_MGE, Inh_CGE, NonN_Astro_FGF3R, NonN_Endo, NonN_Micro, NonN_Oligo_MBP, NonN_OPC) ~ group, data = methylcc_ctp_pheno.tmp))

  summary(lm(cbind(Exc_upper, Exc_lower, Inh_MGE, Inh_CGE, NonN_Astro_FGF3R, NonN_Endo, NonN_Micro, NonN_Oligo_MBP, NonN_OPC) ~ group + age + sex + Batch + PMI, data = methylcc_ctp_pheno.tmp))

}

```

## Test aggregated sums

```{r}

summary(lm(hseq_Glial ~ group, data = ctp_pheno))
summary(lm(mcc_Glial ~ group, data = ctp_pheno))
summary(lm(h_NeuN_neg ~ group, data = ctp_pheno))

summary(lm(hseq_Glial ~ group + age + sex + Batch + PMI, data = ctp_pheno))
summary(lm(mcc_Glial ~ group + age + sex + Batch + PMI, data = ctp_pheno))
summary(lm(h_NeuN_neg ~ group + age + sex + Batch + PMI, data = ctp_pheno))

```

## Adjust for oligodendrocyte proportion

```{r}

if (cellnum == 7) {

  summary(lm(cbind(hseq_Exc, hseq_Inh, hseq_Astro, hseq_Endo, hseq_Micro, hseq_OPC) ~ group, data = hseq_ctp_pheno_adjoligo.tmp))
  summary(lm(cbind(hseq_Exc, hseq_Inh, hseq_Astro, hseq_Endo, hseq_Micro, hseq_OPC) ~ group + age + sex + Batch + PMI, data = hseq_ctp_pheno_adjoligo.tmp))

} else if (cellnum == 9) {
  
  summary(lm(cbind(Exc_upper, Exc_lower, Inh_MGE, Inh_CGE, NonN_Astro_FGF3R, NonN_Endo, NonN_Micro, NonN_OPC) ~ group, data = methylcc_ctp_pheno_adjoligo.tmp))
  summary(lm(cbind(Exc_upper, Exc_lower, Inh_MGE, Inh_CGE, NonN_Astro_FGF3R, NonN_Endo, NonN_Micro, NonN_OPC) ~ group + age + sex + Batch + PMI, data = methylcc_ctp_pheno_adjoligo.tmp))

}

```

# Model: per-cell-type, compare clr-transformed

```{r}

if (cellnum == 7) {

  summary(lm(cbind(hseq_Exc, hseq_Inh, hseq_Astro, hseq_Endo, hseq_Micro, hseq_Oligo, hseq_OPC) ~ group, data = hseq_ctp_pheno.clr.1e3))
  summary(lm(cbind(hseq_Exc, hseq_Inh, hseq_Astro, hseq_Endo, hseq_Micro, hseq_Oligo, hseq_OPC) ~ group + age + sex + Batch + PMI, data = hseq_ctp_pheno.clr.1e3))

} else if (cellnum == 9) {
  
  summary(lm(cbind(Exc_upper, Exc_lower, Inh_MGE, Inh_CGE, NonN_Astro_FGF3R, NonN_Endo, NonN_Micro, NonN_Oligo_MBP, NonN_OPC) ~ group, data = methylcc_ctp_pheno.clr.1e3))
  summary(lm(cbind(Exc_upper, Exc_lower, Inh_MGE, Inh_CGE, NonN_Astro_FGF3R, NonN_Endo, NonN_Micro, NonN_Oligo_MBP, NonN_OPC) ~ group + age + sex + Batch + PMI, data = methylcc_ctp_pheno.clr.1e3))

}

```

## Adjust for oligodendrocyte proportion

```{r}

if (cellnum == 7) {

  summary(lm(cbind(hseq_Exc, hseq_Inh, hseq_Astro, hseq_Endo, hseq_Micro, hseq_OPC) ~ group, data = hseq_ctp_pheno_adjoligo.clr.1e3))
  summary(lm(cbind(hseq_Exc, hseq_Inh, hseq_Astro, hseq_Endo, hseq_Micro, hseq_OPC) ~ group + age + sex + Batch + PMI, data = hseq_ctp_pheno_adjoligo.clr.1e3))

} else if (cellnum == 9) {
  
  summary(lm(cbind(Exc_upper, Exc_lower, Inh_MGE, Inh_CGE, NonN_Astro_FGF3R, NonN_Endo, NonN_Micro, NonN_OPC) ~ group, data = methylcc_ctp_pheno_adjoligo.clr.1e3))
  summary(lm(cbind(Exc_upper, Exc_lower, Inh_MGE, Inh_CGE, NonN_Astro_FGF3R, NonN_Endo, NonN_Micro, NonN_OPC) ~ group + age + sex + Batch + PMI, data = methylcc_ctp_pheno_adjoligo.clr.1e3))

}

```

# Model: per-cell-type, compare ratios between other CTPs and Inh

- dealing with this directly as a ratio, so no need to do transform (what an alr would do anyway)

```{r}

hseq_ctp_pheno.tmp.clr0_ratio <- data.frame(hseq_ctp_pheno.tmp.clr0)
hseq_ctp_pheno.tmp.clr0_ratio$exc_inh <- hseq_ctp_pheno.tmp.clr0_ratio$hseq_Exc/hseq_ctp_pheno.tmp.clr0_ratio$hseq_Inh

hseq_ctp_pheno.tmp.clr0_ratio.df <- data.frame(keep_pheno_col, hseq_ctp_pheno.tmp.clr0_ratio)

summary(lm(exc_inh ~ group, data = hseq_ctp_pheno.tmp.clr0_ratio.df))

summary(lm(exc_inh ~ group + age + sex + Batch + BrainBank, data = hseq_ctp_pheno.tmp.clr0_ratio.df))

```

# Model: ALDEx2 for differential proportion

- Documentation: https://www.bioconductor.org/packages/release/bioc/vignettes/ALDEx2/inst/doc/ALDEx2_vignette.pdf
- Uses Bayesian sampling from a Dirichlet distribution to estimate underlying technical variation
- Input count data, takes as given (requires integers, so rounded the back-derived count data)
- --> infers technical variation (eg. sequencing the same sample again) multiple times using aldex.clr()
    - aldex.clr(): recommend mc.samples >128 for t-test, >1000 for rigorous effect size calculations, >16 for ANOVA 
- Outputs (see p10 of documentation):
    - Tables
        - we*: Welch's t-test
        - wi*: Wilcoxon rank test
        - *ep: expected p-value
        - *eBH: expected BH-correction
        - rab*: median clr value for a given group
        - dif.btw: median difference in clr value between groups
        - dif.win: median of largest difference in clr values within groups
        - effect: median effect size (diff.btw/max(diff.win))
        - overlap: prop of effect size that overlaps 0 (ie. no effect)
    - Plots 
        - relationships between Abundance (clr) vs Difference + Dispersion vs Difference
        - red denotes differentially abundant with q < 0.1, grey abundant but not differentially-abundant, black are rare but not differentially abundant)

## houseman_seq

### No covariates

```{r}

conds <- hseq_ctp_pheno.clr.1e3$group

aldex_in <- sapply(data.frame(t(hseq_ctp_pheno.tmp[,c(all_of(level2_cols))])*10000), as.integer)
rownames(aldex_in) <- level2_cols
colnames(aldex_in) <- hseq_ctp_pheno.tmp$IID

x <- aldex.clr(aldex_in, conds, mc.samples=128, denom="all", verbose=F)

x.tt <- aldex.ttest(x, paired.test = F)

x.effect <- aldex.effect(x, verbose = F)

x.all <- data.frame(x.tt, x.effect)

tmp <- rownames(x.all)
x.all <- as.data.frame(sapply(x.all, function(x) signif(x, 2)))
rownames(x.all) <- tmp

datatable(x.all)

par(mfrow=c(1,2))
aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",ylab="Difference")
aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",ylab="Difference")

```


### glm + covariates

- Runs a glm
- Input model matrix

```{r, message = F}

conds <- hseq_ctp_pheno.tmp$group
cov.aldex <- methylcc_ctp_pheno.tmp %>% dplyr::select(c("group", "age", "sex", "Batch", "BrainBank"))
mm <- model.matrix(~., cov.aldex)

x <- aldex.clr(aldex_in, mm, mc.samples=128, denom="all", verbose=F)

x.glm <- aldex.glm(x, mm)

x.all2 <- data.frame(x.glm)

tmp <- rownames(x.all2)
x.all2 <- as.data.frame(sapply(x.all2, function(x) signif(x, 2)))
rownames(x.all2) <- tmp

datatable(x.all2)

```

# Model: ANCOM for differential proportion

## methylCC

### Prep

```{r}

# Prepare data
metadata <- hseq_ctp_pheno.tmp %>% dplyr::select(c("IID", "group", "age", "sex", "Batch", "BrainBank"))
rownames(metadata) <- metadata$IID
metadata$group <- as.factor(metadata$group)

ancom_features <- t(hseq_ctp_pheno.tmp[,c(level2_cols)])*10000
colnames(ancom_features) <- hseq_ctp_pheno.tmp$IID
ancom_features <- ancom_features[,which(colnames(ancom_features) %in% metadata$IID)]

# Check variables in the correct order
identical(colnames(ancom_features), metadata$IID)

ancom_table <- feature_table_pre_process(ancom_features, metadata, "IID", group_var = NULL, out_cut = 0.05, zero_cut = 0.90, lib_cut = 10, neg_lb)

```

### Case-control analysis with no covariates

```{r}

# Run tests with covariates (age + sex + diet)
ancom_out_nocov <- ANCOM(ancom_table$feature_table, ancom_table$meta_data, ancom_table$structure_zeros, main_var = "group", p_adj_method = "BH")

ancom_out_nocov$out <- ancom_out_nocov$out[order(ancom_out_nocov$out$W, decreasing = T),]

datatable(ancom_out_nocov$out)

ancom_out_nocov$fig

```

### Case-control analysis with covariates

```{r}

# Run tests with covariates (age + sex + diet)
ancom_out_cov <- ANCOM(ancom_table$feature_table, ancom_table$meta_data, ancom_table$structure_zeros, main_var = "group", p_adj_method = "BH", adj_formula = "age + sex + Batch + BrainBank")

ancom_out_cov$out <- ancom_out_cov$out[order(ancom_out_cov$out$W, decreasing = T),]

datatable(ancom_out_cov$out)

ancom_out_cov$fig

```

# Correlation plots

```{r}

cov_var <- c("IID", "group", "age", "sex", "Batch", "BrainBank")

```

## Glial

### Aggregated

- eBayes + Houseman (coefs_brain): h_NeuN_neg
- methylCC (summed): Glial
- smartSVA: sSV1
- eBayes + Houseman (summed coefs_sub): h_Glial
- + include methylCC oligo since it seems to go together ...

```{r}

neunn <- ctp_pheno %>% dplyr::select(c(all_of(cov_var), "h_NeuN_neg", "hseq_Glial", "mcc_Glial", "sSV1", "sSV2", "h_Glial", "hseq_Oligo", "hseq_OPC", "mcc_Oligo", "mcc_OPC", "h_OLIGO", "h_OPC", "celfie_Glial", "celfie_Oligo", "celfie_OPC", "VAEe3", "VAEe9"))

# y = h_NeuN_neg
neunn.long <- melt(neunn, id.var = c(all_of(cov_var), "h_NeuN_neg"), measure.var = c("hseq_Glial", "mcc_Glial", "h_Glial", "celfie_Glial", "sSV1", "VAEe3"), variable.name = c("method"), value.name = "CTP")
ggplot(neunn.long, aes(x=CTP, y=h_NeuN_neg)) + geom_point(aes(colour = age)) + scale_colour_viridis(discrete = FALSE) + geom_smooth(method = "lm") + facet_wrap(~method, scales = "free_x")

# y: hseq_Glial
neunn.long <- melt(neunn, id.var = c(all_of(cov_var), "hseq_Glial"), measure.var = c("h_NeuN_neg", "mcc_Glial", "h_Glial", "celfie_Glial", "sSV1", "VAEe3"), variable.name = c("method"), value.name = "CTP")
ggplot(neunn.long, aes(x=CTP, y=hseq_Glial)) + geom_point(aes(colour = age)) + scale_colour_viridis(discrete = FALSE) + geom_smooth(method = "lm") + facet_wrap(~method, scales = "free_x")

# y: mcc_Glial
neunn.long <- melt(neunn, id.var = c(all_of(cov_var), "mcc_Glial"), measure.var = c("h_NeuN_neg", "hseq_Glial", "h_Glial", "celfie_Glial", "sSV1", "VAEe3"), variable.name = c("method"), value.name = "CTP")
ggplot(neunn.long, aes(x=CTP, y=mcc_Glial)) + geom_point(aes(colour = age)) + scale_colour_viridis(discrete = FALSE) + geom_smooth(method = "lm") + facet_wrap(~method, scales = "free_x")

# y: h_Glial
neunn.long <- melt(neunn, id.var = c(all_of(cov_var), "h_Glial"), measure.var = c("h_NeuN_neg", "hseq_Glial", "mcc_Glial", "celfie_Glial", "sSV1", "VAEe3"), variable.name = c("method"), value.name = "CTP")
ggplot(neunn.long, aes(x=CTP, y=h_Glial)) + geom_point(aes(colour = age)) + scale_colour_viridis(discrete = FALSE) + geom_smooth(method = "lm") + facet_wrap(~method, scales = "free_x")

# y: celfie_Glial
neunn.long <- melt(neunn, id.var = c(all_of(cov_var), "celfie_Glial"), measure.var = c("h_NeuN_neg", "hseq_Glial", "mcc_Glial", "h_Glial", "sSV1", "VAEe3"), variable.name = c("method"), value.name = "CTP")
ggplot(neunn.long, aes(x=CTP, y=celfie_Glial)) + geom_point(aes(colour = age)) + scale_colour_viridis(discrete = FALSE) + geom_smooth(method = "lm") + facet_wrap(~method, scales = "free_x")

# y: sSV1
neunn.long <- melt(neunn, id.var = c(all_of(cov_var), "sSV1"), measure.var = c("h_NeuN_neg", "hseq_Oligo", "mcc_Oligo", "h_OLIGO", "celfie_Oligo", "VAEe9"), variable.name = c("method"), value.name = "CTP")
ggplot(neunn.long, aes(x=CTP, y=sSV1)) + geom_point(aes(colour = age)) + scale_colour_viridis(discrete = FALSE) + geom_smooth(method = "lm") + facet_wrap(~method, scales = "free_x")

# y: sSV1
neunn.long <- melt(neunn, id.var = c(all_of(cov_var), "sSV1"), measure.var = c("h_NeuN_neg", "hseq_Glial", "mcc_Glial", "h_Glial", "celfie_Glial", "hseq_Oligo", "hseq_OPC", "mcc_Oligo", "mcc_OPC", "h_OLIGO", "h_OPC", "celfie_Oligo", "celfie_OPC", "VAEe9"), variable.name = c("method"), value.name = "CTP")
ggplot(neunn.long, aes(x=CTP, y=sSV1)) + geom_point(aes(colour = age)) + scale_colour_viridis(discrete = FALSE) + geom_smooth(method = "lm") + facet_wrap(~method, scales = "free_x")

# y: VAEe9
neunn.long <- melt(neunn, id.var = c(all_of(cov_var), "VAEe9"), measure.var = c("h_NeuN_neg", "hseq_Oligo", "mcc_Oligo", "h_OLIGO", "celfie_Oligo", "sSV1"), variable.name = c("method"), value.name = "CTP")
ggplot(neunn.long, aes(x=CTP, y=VAEe9)) + geom_point(aes(colour = age)) + scale_colour_viridis(discrete = FALSE) + geom_smooth(method = "lm") + facet_wrap(~method, scales = "free_x")

```

### Compare individual glial sub-type CTPs

```{r, fig.height = 10, fig.width = 8, message = F, warning = F}

# Comparison between houseman extremes vs eBayes
ctp_pheno_sub_glial <- data.frame(ctp_pheno[,c("hseq_Glial", "hseq_Astro", "hseq_Endo", "hseq_Micro", "hseq_Oligo", "hseq_OPC", "h_NeuN_neg", "h_Glial", "h_ASTRO", "h_ENDO", "h_MONO", "h_OLIGO", "h_OPC")])

ggpairs(ctp_pheno_sub_glial)

# Comparison between houseman extremes vs methylCC extremes
ctp_pheno_sub_glial <- data.frame(ctp_pheno[,c("hseq_Glial", "hseq_Astro", "hseq_Endo", "hseq_Micro", "hseq_Oligo", "hseq_OPC", "h_NeuN_neg", "h_Glial", "h_ASTRO", "h_ENDO", "h_MONO", "h_OLIGO", "h_OPC")])

ggpairs(ctp_pheno_sub_glial)

# Comparison between houseman extremes vs CelFIE
ctp_pheno_sub_glial <- data.frame(ctp_pheno[,c("hseq_Glial", "hseq_Astro", "hseq_Endo", "hseq_Micro", "hseq_Oligo", "hseq_OPC", "h_NeuN_neg", "celfie_Glial", "celfie_Astro", "celfie_Endo", "celfie_Micro", "celfie_Oligo", "celfie_OPC")])

ggpairs(ctp_pheno_sub_glial)

# Comparison between houseman array vs methylCC extremes
ctp_pheno_sub_glial <- data.frame(ctp_pheno[,c("h_Glial", "h_ASTRO", "h_ENDO", "h_MONO", "h_OLIGO", "h_OPC", "h_NeuN_neg", "mcc_Glial", "mcc_Astro", "mcc_Endo", "mcc_Micro", "mcc_Oligo", "mcc_OPC")])

ggpairs(ctp_pheno_sub_glial)

```

## Neuronal 

### Aggregated

```{r}

# Select columns
neunp <- ctp_pheno %>% dplyr::select(c(all_of(cov_var), "h_NeuN_pos", "hseq_Neuron", "mcc_Neuron", "h_Neuron", "celfie_Neuron", "sSV1", "VAEe3"))

# y: h_NeuN_pos
neunp.long <- melt(neunp, id.var = c(all_of(cov_var), "h_NeuN_pos"), measure.var = c("hseq_Neuron", "mcc_Neuron", "h_Neuron", "celfie_Neuron", "sSV1", "VAEe3"), variable.name = c("method"), value.name = "CTP")
ggplot(neunp.long, aes(x=CTP, y=h_NeuN_pos)) + geom_point(aes(colour = age)) + scale_colour_viridis(discrete = F) + geom_smooth(method = "lm") + facet_wrap(~method, scales = "free_x")

# y: hseq_Neuron
neunp.long <- melt(neunp, id.var = c(all_of(cov_var), "hseq_Neuron"), measure.var = c("h_NeuN_pos", "mcc_Neuron", "h_Neuron", "celfie_Neuron", "sSV1", "VAEe3"), variable.name = c("method"), value.name = "CTP")
ggplot(neunp.long, aes(x=CTP, y=hseq_Neuron)) + geom_point(aes(colour = age)) + scale_colour_viridis(discrete = F) + geom_smooth(method = "lm") + facet_wrap(~method, scales = "free_x")

# y: mcc_Neuron
neunp.long <- melt(neunp, id.var = c(all_of(cov_var), "mcc_Neuron"), measure.var = c("h_NeuN_pos", "hseq_Neuron", "h_Neuron", "celfie_Neuron", "sSV1", "VAEe3"), variable.name = c("method"), value.name = "CTP")
ggplot(neunp.long, aes(x=CTP, y=mcc_Neuron)) + geom_point(aes(colour = age)) + scale_colour_viridis(discrete = F) + geom_smooth(method = "lm") + facet_wrap(~method, scales = "free_x")

# y: h_Neuron
neunp.long <- melt(neunp, id.var = c(all_of(cov_var), "h_Neuron"), measure.var = c("h_NeuN_pos", "hseq_Neuron", "mcc_Neuron", "celfie_Neuron", "sSV1", "VAEe3"), variable.name = c("method"), value.name = "CTP")
ggplot(neunp.long, aes(x=CTP, y=h_Neuron)) + geom_point(aes(colour = age)) + scale_colour_viridis(discrete = F) + geom_smooth(method = "lm") + facet_wrap(~method, scales = "free_x")

# y: celfie_Neuron
neunp.long <- melt(neunp, id.var = c(all_of(cov_var), "celfie_Neuron"), measure.var = c("h_NeuN_pos", "hseq_Neuron", "mcc_Neuron", "h_Neuron", "sSV1", "VAEe3"), variable.name = c("method"), value.name = "CTP")
ggplot(neunp.long, aes(x=CTP, y=celfie_Neuron)) + geom_point(aes(colour = age)) + scale_colour_viridis(discrete = F) + geom_smooth(method = "lm") + facet_wrap(~method, scales = "free_x")

# y: sSV1
neunp.long <- melt(neunp, id.var = c(all_of(cov_var), "sSV1"), measure.var = c("h_NeuN_pos", "hseq_Neuron", "mcc_Neuron", "h_Neuron", "celfie_Neuron", "VAEe3"), variable.name = c("method"), value.name = "CTP")
ggplot(neunp.long, aes(x=CTP, y=sSV1)) + geom_point(aes(colour = age)) + scale_colour_viridis(discrete = F) + geom_smooth(method = "lm") + facet_wrap(~method, scales = "free_x")

# y: VAEe3
neunp.long <- melt(neunp, id.var = c(all_of(cov_var), "VAEe3"), measure.var = c("h_NeuN_pos", "hseq_Neuron", "mcc_Neuron", "h_Neuron", "celfie_Neuron", "sSV1"), variable.name = c("method"), value.name = "CTP")
ggplot(neunp.long, aes(x=CTP, y=VAEe3)) + geom_point(aes(colour = age)) + scale_colour_viridis(discrete = F) + geom_smooth(method = "lm") + facet_wrap(~method, scales = "free_x")

```

### Compare individual neuronal sub-type CTPs

```{r, fig.height = 10, fig.width = 8, message = F, warning = F}

if (cellnum == 7) {
  ctp_pheno_sub_neuron <- data.frame(ctp_pheno[,c("hseq_Neuron", "hseq_Exc", "hseq_Inh", "h_Neuron", "h_GLU", "h_GABA", "mcc_Neuron", "mcc_Exc", "mcc_Inh", "h_NeuN_pos")])

} else if (cellnum == 9) {
  ctp_pheno_sub_neuron <- data.frame(ctp_pheno[,c("Neuron", "Exc_upper", "Exc_lower", "Inh_MGE", "Inh_CGE", "h_NeuN_pos", "h_Neuron", "h_GABA", "h_GLU")])
}

ggpairs(ctp_pheno_sub_neuron)

```

## All together

```{r}

ctp_pheno_sub <- data.frame(ctp_pheno[,c("hseq_Exc", "hseq_Inh", "hseq_Astro", "hseq_Endo", "hseq_Micro", "hseq_Oligo", "hseq_OPC", "h_GLU", "h_GABA", "h_ASTRO", "h_ENDO", "h_MONO", "h_OLIGO", "h_OPC")])

#svg(paste(out_dir, "/", filen, "_ctp_hseq_harray_cor_matrix.svg", sep = ""), height = 6, width = 6)
jpeg(paste(out_dir, "/", filen, "_ctp_hseq_harray_cor_matrix.jpeg", sep = ""), height = 600, width = 600, quality = 100)
g <- ggduo(ctp_pheno_sub, colnames(ctp_pheno_sub)[1:7], colnames(ctp_pheno_sub)[8:14], xlab = "Houseman + sequencing reference", ylab = "Houseman + array reference", title = "UCLA_ASD")
print(g)
dev.off()

```

## Unsupervised comparison

### Neurons

```{r, fig.height = 10, fig.width = 8, message = F, warning = F}

if (cellnum == 7) {
  ctp_pheno_sv_neuron <- data.frame(ctp_pheno[,c("h_NeuN_pos", "hseq_Neuron", "hseq_Exc", "hseq_Inh", "sSV1", "sSV2", "sSV3", "sSV4", "sSV5", "sSV6", "sSV7", "sSV8")])
} else if (cellnum == 9) {
  ctp_pheno_sv_neuron <- data.frame(ctp_pheno[,c("h_NeuN_pos", "Neuron", "Exc_upper", "Exc_lower", "Inh_MGE", "Inh_CGE", "sSV1", "sSV2", "sSV3", "sSV4", "sSV5", "sSV6", "sSV7", "sSV8")])
}

ggpairs(ctp_pheno_sv_neuron)

```

### Glial

```{r, fig.height = 10, fig.width = 8, message = F, warning = F}

ctp_pheno_sv_glial <- data.frame(ctp_pheno[,c("h_NeuN_neg", "hseq_Glial", "hseq_Astro", "hseq_Endo", "hseq_Micro", "hseq_Oligo", "hseq_OPC", "sSV1", "sSV2", "sSV3", "sSV4", "sSV5", "sSV6", "sSV7", "sSV8")])

ggpairs(ctp_pheno_sv_glial)

```

# For presentations

```{r, fig.height = 6, fig.width = 6}

# For presentations
ggplot(neunn, aes(x = sSV1, y = hseq_Oligo)) + geom_point() + geom_smooth(method = "lm") + xlab("smartSV1 (reference-free)") + ylab("Houseman (seq) Oligo proportion")
ggsave(paste(out_dir, "/hseqOligo_vs_smartSV1.png", sep = ""), height = 4, width = 4)
ggplot(neunn, aes(x = h_NeuN_neg, y = hseq_Glial)) + geom_point() + geom_smooth(method = "lm") + xlab("NeuN_neg proportion (reference-based)") + ylab("Sum of Houseman (seq) Glial proportions")
ggsave(paste(out_dir, "/NeuN_neg_vs_hseqGlial.png", sep = ""), height = 4, width = 4)

```